Alien species

### 2023-03-22 DRAFT_alien_species.Rmd
latitudinal effect

>>>>

#### Effect of latitude

Alien species tend to invade from the south. Let us see if there is an effect of latitude.

Extracting x and y coordinates

```{r}
temp <- alien_extr %>%
  filter(!is.na(w_mean)) %>%
  st_centroid() %>%
  st_coordinates()

temp2 <- alien_extr %>%
  filter(!is.na(w_mean))  %>%
  cbind(temp)
```

```{r lat-effect, fig.cap="Effect of latitude on the alien species indicator values. The blue line is a loess smoother (span=0.75)."}
temp2 %>%
  ggplot(aes(x = X, y = w_mean))+
  geom_point()+
  theme_bw()+
  geom_smooth(linewidth=2,
              method="loess",
              span=0.75)+
  labs(x = "Latitude (UTM with offset)",
       y = "Alien species indicator value\n(area weighted)")+
  facet_wrap(~eco)

```{r, eval=F, include=F}
# Here is the same analysis, but at the polygon scale (it's a bit more messy)
temp <- alien_data %>%
  st_centroid() %>%
  st_coordinates()

temp2 <- alien_data %>%
  cbind(temp)

temp2 %>%
  ggplot(aes(x = X, y = i_2020))+
  geom_point()+
  theme_bw()+
  geom_smooth(linewidth=2,
              method="loess",
              span=0.75)+
  labs(x = "Latitude (EPSG:25833)",
       y = "Alien species indicator value")+
  facet_wrap(~hovedøkosystem)

There is a quite clear association here between latitude and the on-site effect of alien species.
